---
title: "Assignment 3"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Packages, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

library(tidyverse)
library(stargazer)
#load csv
 
hw3 <- read_csv("HW3Data.csv")


```


## Part 1. Analyzing costs and benefits of a VSR program

<br>

###1. Linear Probability Model 

```{r Question 1, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

# Create a linear probability model that predicts a respondent's probability of voting 'yes' on the ballot based on their age, income, NEP score, the risk reduction offered by the program (risk), and the cost of the program to that respondent (bid). 

lm_vote <- lm(formula = vote ~ risk + bid + NEP + income + age, data = hw3)

summary(lm_vote)



# Coefficients:
#       (Vote)               risk                bid                NEP     incomeone_percent         incomepoor  
#         0.1196977          0.0007445         -0.0010699          0.0158639          0.0088282          0.0027386  

#        incomerich    incomevery_rich         agetofifty         agetoforty         agetosixty        agetothirty  
#         0.0074891          0.0467922          0.0099816         -0.0201190         -0.0162261          0.0204401 



# Analyzing coefficients:


# For every 1% increase in risk reduction, there is an increase in probablility of voting YES by 0.0007445.
# For every $1 increase in cost per household, there is a decrease in probability of voting YES by 0.0010699.
# For every additional point increase for NEP score, there is an increase in probability of voting YES by 0.0158639.
# Individuals in the top one percent income category are 0.0088282 more likely to vote YES than individuals in the middle income class.
# Individuals in the poor income category are 0.0027386 more likely to vote YES than inidividuals in the middle income category. 
# Individuals in the rich income category are 0.0074891 more likely to vote YES than inidividuals in the middle income category.
# Individuals in the very rich income category are 0.0467922 more likely to vote YES than inidividuals in the middle income category.

# Age to fifty individuals are 0.0099816 more likely to vote YES than age over 60.
# Age to 40 individuals are 0.0201190 more likely to vote YES than age over 60.
# Age to 60 are 0.0162261 are less likely to vote YES than age over 60.
# Age to 30 are 0.0204401 are less likely to vote YES than age over 60.


#assign the variables

vote_int <- lm_vote$coefficients[1]
risk <- lm_vote$coefficients[2]
bid <- lm_vote$coefficients[3]
NEP <- lm_vote$coefficients[4]
inc_1 <- lm_vote$coefficients[5]
inc_poor <- lm_vote$coefficients[6]
inc_rich <- lm_vote$coefficients[7]
inc_vrich <- lm_vote$coefficients[8]
age_50 <- lm_vote$coefficients[9]
age_40 <- lm_vote$coefficients[10]
age_60 <- lm_vote$coefficients[11]
age_30 <- lm_vote$coefficients[12]


# Put linear model into stargazer table 

# stargazer(lm_vote)



```

$Probability(Voting~Yes) = `r round(vote_int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) + `r round(age_60, digits=4)`(Age~to~60) + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(inc_vrich,digits=4)`(Income~Very~Rich) + `r round(NEP, digits=4)`(NEP) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

Interpreting the coefficients:

a. Age to 30 are 0.0204401 are less likely to vote YES than age over 60.
b. Age to 40 individuals are 0.0201190 more likely to vote YES than age over 60.
c. Age to 50 individuals are 0.0099816 more likely to vote YES than age over 60.
d. Age to 60 are 0.0162261 are less likely to vote YES than age over 60.
e. Individuals in the top one percent income category are 0.0088282 more likely to vote YES than individuals in the middle income class.
f. Individuals in the poor income category are 0.0027386 more likely to vote YES than inidividuals in the middle income category.
g. Individuals in the rich income category are 0.0074891 more likely to vote YES than inidividuals in the middle income category.
h. Individuals in the very rich income category are 0.0467922 more likely to vote YES than inidividuals in the middle income category.
i. For every additional point increase for NEP score, there is an increase in probability of voting YES by 0.0158639.
j. For every $1 increase in annual payment for a household (bid), there is a decrease in probability of voting YES by 0.0010699.
k. For every 1% increase in risk reduction, there is an increase in probablility of voting YES by 0.0007445.

<br>
<br>

###2. Value of a single prevented whale death 

```{r Question 2, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

## Based on this regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)

#In order to reduce the risk of one whale strike, reduce the risk to 4%. 

#The average of voting YES is 0.5

#Compare the willingness to pay at business as usual (VSR of 0) to the willingness to pay when increasing the VSR to 4%. 

#average of voting yes or no (0 or 1) is 0.5
avg_vote <- 0.5



#average of NEP
NEP_avg <- mean(hw3$NEP)  


#Find the Bid (cost to households) when reduction is 0% = 269 

bid_zero <- (vote_int + (risk*0) + (NEP*NEP_avg) + inc_1 + inc_poor + inc_vrich + inc_rich + age_50 + age_30 + age_40 + age_60 - avg_vote)/(-bid)

#Find the Bid (cost to households) when reduction is 4% = 272

bid_four <- (vote_int + (risk*4) + (NEP*NEP_avg) + inc_1 + inc_poor + inc_rich + inc_vrich + age_50 + age_30 + age_40 + age_60 - avg_vote)/(-bid)

#The value of a single whale is the difference between the willingness to pay (i.e.Bid) for a vessel speed reduction program at 4% and at 0%. = 2.78

wtp_value <- bid_four - bid_zero 


```

**The Bid when reduction is 0%. The average vote between 0 (no) and 1 (yes) is 0.5. **
$0.5 = `r round(vote_int, digits = 4)` + `r round(risk, digits=4)`(0) + `r round(NEP, digits=4)`(NEP) + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(inc_vrich, digits=4)`(Income~Very~Rich)  + `r round(inc_rich, digits=4)`(Income~Rich) +`r round(age_50, digits=4)`(Age~to~50) +`r round(age_30, digits=4)`(Age~to~30) +`r round(age_40, digits=4)`(Age~to~40) +`r round(age_60, digits=4)`(Age~to~60) + `r round(bid, digits=4)`(Bid)$

$Willingness~to~Pay = `r round(bid_zero, digits=2)`$

<br>
**The Bid when reduction is 4%. **
$0.5 = `r round(vote_int, digits = 4)` + `r round(risk, digits=4)`(4) + `r round(NEP, digits=4)`(NEP) + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(inc_vrich, digits=4)`(Income~Very~Rich)  + `r round(inc_rich, digits=4)`(Income~Rich) +`r round(age_50, digits=4)`(Age~to~50) +`r round(age_30, digits=4)`(Age~to~30) +`r round(age_40, digits=4)`(Age~to~40) +`r round(age_60, digits=4)`(Age~to~60) + `r round(bid, digits=4)`(Bid)$

$Willingness~to~Pay = `r round(bid_four, digits=2)`$ 

<br>

**The value of a single whale is the difference between the willingness to pay (i.e.Bid) for a vessel speed reduction program at 4% and at 0%. **

$Individual~Whale~Value = `r round(wtp_value, digits=2)`$ 


<br>
<br>

### 3. Estimated Willingness to Pay for three arbitrary respondents for 60% risk reduction VSR program.


```{r Question 3, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

# Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR program offering 60% risk reduction


# Filter for risk = 60%

risk_60 <- hw3 %>% 
  filter(risk == "60")

# Respondent 8:
# Risk = 60
# Bid = 80
# Vote = 1
# NEP = 29
# Income = Very_Rich
# Age = ToThirty
# VSR = 60%

bid_8 <-(vote_int + (NEP*29) + inc_vrich + age_30 - avg_vote)/(-bid) 
# bid_8 = $137


# Respondent 63:
# Risk = 60
# Bid = 20
# Vote = 1
# NEP = 37
# Income = Rich
# Age = OverSixty

bid_63 <-(vote_int + (NEP*37) + inc_rich - avg_vote)/(-bid) 
#bid_63 = $200


# Respondent 7:
# Risk = 60
# Bid = 40
# Vote = 0
# NEP = 22
# Income = One Percent
# Age = ToForty

bid_7 <-(vote_int + (NEP*22) + inc_1 + age_40 - avg_vote)/(bid) 
#bid_7 = $39.80


```


**Three arbitrary respondents:**

- 8   NEP: 29   Income: Very Rich   Age: To Thirty    Vote: 1

- 63  NEP: 37   Income: Rich        Age: Over Sixty   Vote: 1

- 7   NEP: 22   Income: One Percent Age: To Forty     Vote: 0

<br>

**Willingness to Pay for 60% risk reduction program: **

Individual 8: 

$`r avg_vote` = `r round(vote_int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_vrich, digits=4)`(Income~Very~Rich) + `r round(NEP, digits=4)`(NEP*29) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

$Willingness~to~Pay = `r round(abs(bid_8), digits=2)`$    

<br>

Individual 63: 

$`r avg_vote` = `r round(vote_int, digits = 4)` + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(NEP, digits=4)`(NEP*37) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

$Willingness~to~Pay = `r round(abs(bid_63), digits=2)`$    

<br>


Individual 7:

$`r avg_vote` = `r round(vote_int, digits = 4)` + `r round(age_40, digits=4)`(Age~to~40) + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(NEP, digits=4)`(NEP*22) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

$Willingness~to~Pay = `r round(abs(bid_7), digits=2)`$    

<br>
<br>

### 4. Willingness to Pay for 60% risk reduction VSR program among average Santa Barbara households. 
```{r Question 4, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

#find mode of the categorical values: income : "one_percent"
income_mode <- hw3$income

income_mode[hw3 %>%
                match(income_mode) %>% 
                tabulate() %>%
                which.max()]

##find mode of the categorical values: age ; "tothirty"

age_mode <- hw3$age

age_mode[hw3 %>%
                match(age_mode) %>% 
                tabulate() %>%
                which.max()]

#Again assume the probability of voting yes the average of voting yes ($p = .5). Use the most common income bracket (very rich), the most common age bracket (to 30), and average NEP (38.366) to calculate willingness to pay using:

bid_sb <- (vote_int + (NEP*NEP_avg) + age_30 + inc_1 + (risk*60) - avg_vote)/(-bid)

#the average WTP for all of Santa Barbara residents for 60% risk reduction program is $283


```

The probability of voting yes with the average of a yes vote ($p = 0.5$). We will use the mode income (one percent), the mode age (to 30), and average NEP (38.366) to calculate willingness to pay using:

$`r avg_vote` = `r round(vote_int, digits = 4)` + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(age_30, digits=4)`(Age~to~30) + `r round(NEP, digits=4)`(NEP) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

<br>

$Average~Santa~Barbara~County~Household~Willingness~to~Pay = `r round(abs(bid_sb), digits=2)`$ 

<br>
<br>


### 5. Total Benefit of a VSR program for 150,000 households in Santa Barbara County
```{r Question 5, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

# If there are 150,000 households in SB County, estimate the total benefit of a VSR program to these households

bid_sbtotal <- ((vote_int + (NEP*NEP_avg) + age_30 + inc_1 + (risk*60) - avg_vote)/(-bid))* 150000

# The total benefit of a VSR program for all 150,000 [average] households is $42,378,337


```
Calculate the Total Benefits using the following equation: 

$`r avg_vote` = (`r round(vote_int, digits = 4)` + `r round(inc_1, digits=4)`(Income~One~Percent) + `r round(age_30, digits=4)`(Age~to~30) + `r round(NEP, digits=4)`(NEP) + `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk))*(150000)$

<br>

$Total~Benefit = `r format(bid_sbtotal, scientific = F) `$ 

<br>
<br>


### 6. Do the benefits of the VSR program outweigh the costs? 

```{r Question 6, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

# Based on the SB Total benefits, do the benefits of the VSR program outweigh the costs?

# Costs of VSR program = $7 million
# Total benefits to SB households = $42,378,337

# YES, benefits outweigh the costs!

```
- Costs of VSR program = $7 million
- Total benefits to Santa Barbara households = $`r format(bid_sbtotal, scientific = F) ` 

<br>
$The~benefits~outwiegh~the~costs$

<br>
<br>

## Part 2. Evaluating the costs and benefits of carbon trading markets to mitigate fatal whale strikes

<br>


### 7. At what price will the shipping industry decide to adopt the VSR for purely self-interested reasons? 

```{r Question 7, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

#The shipping industry wants the $Z per ton of carbon credits to equal the cost of 60% risk reduction at $1000. The amount of fewer emitted tons (20) multiplied by the price per ton should equal $1000. 
Z <- 1000/20


```

The shipping industry wants the price per ton $(\$Z)$ of carbon credits to equal the cost of 60% risk reduction at $1000. The amount of fewer emitted tons $(20)$ multiplied by the price per ton $(\$Z)$ should equal $1000.

$20*Z=1000$

$Z = \$`r Z`~per~ton$

<br>
<br>

### 8a. Approximately how many whales would be saved if all ships voluntarily adopted the VSR?
```{r Question 8a, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE}

# The exact Value of the carbon credit is $50 per ton, and all ships voluntarily reduce speed to acheive 60% risk reduction. How many whales will be saved? 

# A 20% risk reduction is expected to save at least 5 whales
# Therefore, a 60% risk reduction would be 5*3 = 15 whales saved

```

If 5 whales are saved with a 20% risk reduction program, then a 60% risk reduction program would save aproximately 3 times that amount of whales, equaling 15 whales saved. 

<br>

### 8b. What would be the social value of allowing ships to enter the carbon trading market? 

```{r Question 8b, echo = FALSE, include = FALSE, message = FALSE, warning = FALSE }


# The social value of the carbon trading market would be the total benefits minus the costs of the program. If the ships are voluntarily reducing speed, then the $7 million costs are avoided. Therefore, the social value of allowing ships to enter the carbon trading market will be the Total Benefit calculated from problem 5.  

# Social value = $41,542,642




```

The social value of the carbon trading market would be the total benefits minus the costs of the program. If the ships are voluntarily reducing speed, then the $7 million costs are avoided. Therefore, the social value of allowing ships to enter the carbon trading market will be the Total Benefit calculated from problem 5.  

<br>
Social value = `r format(bid_sbtotal, scientific = F) `



***
###*EXTRA CREDIT*

```{r}

#Create a logit regression and compare to the linear
log_vote <- glm(formula = vote ~ risk + bid + NEP + income + age, data = hw3, family = binomial)

log_vote
# Coefficients:
#       (Intercept)               risk                bid                NEP  
#         -2.088292           0.004169          -0.005865           0.083213  
# incomeone_percent         incomepoor         incomerich    incomevery_rich  
#          0.059366           0.040298           0.060219           0.277879  
#        agetofifty         agetoforty         agetosixty        agetothirty  
#          0.062771          -0.127045          -0.107886           0.092936  


#The LM Coefficients
# Coefficients:
#       (Vote)               risk                bid                NEP     incomeone_percent         incomepoor  
#         0.1196977          0.0007445         -0.0010699          0.0158639          0.0088282          0.0027386  

#        incomerich    incomevery_rich         agetofifty         agetoforty         agetosixty        agetothirty  
#         0.0074891          0.0467922          0.0099816         -0.0201190         -0.0162261          0.0204401 


logvote_int <- log_vote$coefficients[1]
logrisk <- log_vote$coefficients[2]
logbid <- log_vote$coefficients[3]
logNEP <- log_vote$coefficients[4]
loginc_1 <- log_vote$coefficients[5]
loginc_poor <- log_vote$coefficients[6]
loginc_rich <- log_vote$coefficients[7]
loginc_vrich <- log_vote$coefficients[8]
logage_50 <- log_vote$coefficients[9]
logage_40 <- log_vote$coefficients[10]
logage_60 <- log_vote$coefficients[11]
logage_30 <- log_vote$coefficients[12]



```

## Interpreting the Log Regression coefficients:

a. Age to 30 are 0.092936  are more likely to vote YES than age over 60.
b. Age to 40 individuals are 0.127045 less likely to vote YES than age over 60.
c. Age to 50 individuals are 0.062771 more likely to vote YES than age over 60.
d. Age to 60 are 0.107886 are less likely to vote YES than age over 60.
e. Individuals in the top one percent income category are 0.059366 more likely to vote YES than individuals in the middle income class.
f. Individuals in the poor income category are 0.040298 more likely to vote YES than inidividuals in the middle income category.
g. Individuals in the rich income category are 0.060219 more likely to vote YES than inidividuals in the middle income category.
h. Individuals in the very rich income category are 0.277879  more likely to vote YES than inidividuals in the middle income category.
i. For every additional point increase for NEP score, there is an increase in probability of voting YES by 0.083213.
j. For every $1 increase in annual payment for a household (bid), there is a decrease in probability of voting YES by 0.005865.
k. For every 1% increase in risk reduction, there is an increase in probablility of voting YES by 0.004169.

<br>

## Value of a single whale: 

Value of a single whale changed from \$2.78 to \$2.84 


<br>

<<<<<<< HEAD
## Willingness to pay: 
```{r}

#Find the Bid (cost to households) when reduction is 0% = 164 

bid_zero_log <- (logvote_int + (risk*0) + (logNEP*NEP_avg) + loginc_1 + loginc_poor + loginc_vrich + loginc_rich + logage_50 + logage_30 + logage_40 + logage_60 - avg_vote)/(-logbid)

#Find the Bid (cost to households) when reduction is 4% = 167

bid_four_log <- (logvote_int + (logrisk*4) + (logNEP*NEP_avg) + loginc_1 + loginc_poor + loginc_rich + loginc_vrich + logage_50 + logage_30 + logage_40 + logage_60 - avg_vote)/(-logbid)

#The value of a single whale is the difference between the willingness to pay (i.e.Bid) for a vessel speed reduction program at 4% and at 0%. = 2.84
  
wtp_value_log <- bid_four_log - bid_zero_log 



```

=======
## Willingness to pay: 
>>>>>>> e6685d6aa529f91be17048db0bf1991b89f6ac05



